---
title: "Untitled"
author: "Eric Graziosi"
date: "2026-09-26"
output:
  github_document:
    toc: true
---


```{r}
library(tidyverse)
organs_original<- read_csv("C:/Users/ericg/Downloads/global-organ-donation_2018.csv", 
                            col_types = list(`Utilized DBD` = col_number(),
                                             `DD Lung Tx` = col_number(),
                                             `Total Utilized DD` = col_number(),
                                             `LD Lung Tx` = col_number())) 
```

```{r}
organs_original |>
  group_by(REPORTYEAR) |>
  summarise(
    total_rows = n(),
    non_missing = sum(!is.na(`TOTAL Actual DD`))
  ) |>
  ggplot() +
  geom_col(aes(x = REPORTYEAR, y = total_rows)) +
  geom_col(aes(x = REPORTYEAR, y = non_missing))
```

```{r}
organs_original |>
  group_by(REPORTYEAR) |>
  summarise(
    total_rows = n(),
    missing = sum(is.na(`TOTAL Actual DD`)),
    non_missing = sum(!is.na(`TOTAL Actual DD`))
  )
```

From 2000-2014 there are consistently 194 rows per year, but the number of missing values in TOTAL_ACTUAL_DD values increase over time. After 2014, the number of rows decrease, which can indicate that the data collection changed, making the data less complete. TOTAL_ACTUAL_DD is almost completely missing after 2014, with only 7 non-missing values in 2015 and none in 2016 or 2017. Therefore, the lack of completeness after 2014 is due to both fewer records being available and the absence of TOTAL_ACTUAL_DD. 